---
title: "Final-Project"
author: "Tessa Birt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Final Project Ideas

-   Colorado River map with data

    -   water levels, peak flow, drought severity

-   Lake Mead

    -   water level each year

    -   Drought intensity map

-   USGS data

Map of Lake Mead USGS

```{r}
install.packages("rnaturalearth")
install.packages("sf")

library(dataRetrieval)
library(rnaturalearth)
library(sf)
library(dplyr)

site_number <- "15010005"
pCode <- "00062"
start.date <- "2001-10-01"
end.date <- "2022-09-30"
lakemead_data <- readNWISuv(site_number, parameterCd = pCode, startDate = start.date,
                     endDate = end.date)
```

```{r}
library(shiny)

Year <- c(2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010,
          2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021,
          2022)
mead_annual_peak <- c(1760, 1549, 1388, 1967, 1326, 1765, 1814, 1693, 1502, 1751)
mead_annual_low <- c(2176, 3154, 1138, 1196, 2132, 3176, 4181, 5169, 3150, 4175)
mead_annual_average  <- c(2176, 3154, 1138, 1196, 2132, 3176, 4181, 5169, 3150, 4175)
my_data <- as.data.frame(cbind(Year,Auckland,Wellington, Lyttelton))

ui <- fluidPage(
    titlePanel("New Zealand Annual Mean Sea Level (MSL) Summary"),

    sidebarLayout(
        sidebarPanel(
            helpText("Annual Mean Sea Level Summary for various locations around NZ."),

            selectInput("var", 
                        label = "Choose a Location",
                        choices = c("Auckland",
                                    "Lyttelton",
                                    "Wellington"),
                        selected = "Auckland"),

            sliderInput("range", 
                        label = "Choose a start and end year:",
                        min = min(my_data$Year), max = max(my_data$Year), value = c(2003, 2008),sep = "",)
        ),

        mainPanel(
            tableOutput("DataTable")
        )
    )
)
server <- function(input, output) {

    output$DataTable <- renderTable({
        dt <- my_data[my_data$Year >= input$range[1] & my_data$Year <= input$range[2],]
        dt[,c("Year",input$var)]
    },include.rownames=FALSE)

}
shinyApp(ui, server)
```

Yearly peak and low water levels:

<https://www.usbr.gov/lc/region/g4000/lakemead_line.pdf>

Monthly data:

<https://www.usbr.gov/lc/region/g4000/hourly/mead-elv.html>

```{r}
library(readxl)
LakeMead_Monthly_ <- read_excel("~/Desktop/Environmental Data Sci/Tessa-Final-Project-Explore/LakeMead_Monthly.xlsx")
View(LakeMead_Monthly_)

Mead_monthly_filtered <- LakeMead_Monthly_ %>%
  filter(Year > 2000)

Mean_mead <- Mead_monthly_filtered %>%
 mutate(across(starts_with(c("JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT", "NOV", "DEC")), as.numeric)) %>%
  mutate(yearly_mean = rowMeans(select(., starts_with(c("JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT", "NOV", "DEC"))), na.rm = TRUE))

mead_means <- subset(Mean_mead, select = c(Year, yearly_mean))

#Load in mead high and low 
library(readxl)
Mead_high_low <- read_excel("Mead_high_low.xlsx")
View(Mead_high_low)


```

Join Data

```{r}
joined_mead <- left_join(Mead_high_low, mead_means, by = "Year")
```

Plot high, low, and mean

```{r}
library(tidyr)
library(ggplot2)
mead_long <- clean_mead %>%
  gather(key = "depth_type", value = "depth", -Year)

ggplot(mead_long, aes(x = Year, y = depth, color = depth_type)) +
  geom_line() +
  geom_point() +
  labs(title = "Yearly Water Depth for Lake Mead",
       x = "Year",
       y = "Water Depth",
       color = "Depth Type") +
  theme_minimal()

joined_mead$Mean <- joined_mead$yearly_mean

clean_mead <- subset(joined_mead, select = -yearly_mean)
```

Shiny App Workflow

```{r}
# Define UI for application that draws a histogram
ui <- fluidPage(
  # Application title
  titlePanel("Lake Mead Depth"),
  
  # Sidebar with a slider input for number of bins 
  selectInput("depth_type", "Select Depth Type",
              choices = c("Low", "Mean", "High"),
              selected = "Mean"),
  
  # Show a plot of the generated distribution
  mainPanel(
    plotOutput("depth_plot")
  )
) 

# Define server logic required to draw a histogram
server <- function(input, output) {
  
  selected_data <- reactive({
    filter(mead_long, depth_type == input$depth_type)
  })
  
  output$depth_plot <- renderPlot({
    ggplot(data = selected_data(), aes(x = Year, y = depth)) +
      geom_line() +
      geom_point() +
      labs(title = paste(input$depth_type, "Lake Mead Water Depth"),
           x = "Year", y = "Water Depth")
  })
}

# Run the application 
shinyApp(ui = ui, server = server)
```

Shapefiles

```{r}
library(sf)
library(leaflet)
library(dplyr)

shapefile_2001 <- st_read("USDM_20010102_M/USDM_20010102.shp")

my_map_2001 <- leaflet() %>%
  addTiles() %>%
  addPolygons(data = shapefile_2001, fillOpacity = 0.5)

print(my_map_2001)

shapefile_2022 <- st_read("USDM_20221227_M/USDM_20221227.shp")

my_map_2022 <- leaflet() %>%
  addTiles() %>%
  addPolygons(data = shapefile_2022, fillOpacity = 0.5)

print(my_map_2022)
```

```{r}
shapefile_2001 <- st_read("USDM_20010102_M/USDM_20010102.shp")

bbox22 <- st_bbox(c(xmin = -114.8395, ymin = 36.001, xmax = -114.0988, ymax = 36.4334), crs = st_crs(shapefile_2022))

shapefile <- st_make_valid(shapefile_2022)

cropped_2022 <- st_crop(shapefile, bbox22)

print(cropped_2022)

mymap2022 <- cropped_2022 %>%
  leaflet() %>%
  addTiles() %>%
  addPolygons()

print(mymap2022)
```

Mead depth ggplot and table workflow shiny

```{r}
library(shiny)

# Define UI for application that draws a histogram
ui <- fluidPage(
  titlePanel("Lake Mead Water Depth Analysis"),
  
  sidebarLayout(
    sidebarPanel(
      sliderInput("year", "Select Year:", min = 2001, max = 2022, value = 2001, step = 1, sep = "")
    ),
    
    mainPanel(
      plotOutput("staticPlot"),
      tableOutput("summaryTable")
    )
  )
)

# Define server
server <- function(input, output) {

  # Create ggplot based on filtered data
  static_plot <- ggplot(mead_long, aes(x = Year, y = depth, color = depth_type)) +
    geom_line() +
    labs(title = "Yearly Water Depth for Lake Mead",
         x = "Year",
         y = "Water Depth",
         color = "Depth Type") +
    theme_minimal()
  
  # Display static plot
  output$staticPlot <- renderPlot({
    print(static_plot)
  })
  
  filtered_data <- reactive({
    filter(mead_long, Year == input$year)
  })
  
  # Create summary table based on filtered data
  output$summaryTable <- renderTable({
    summary_data <- filtered_data() %>%
      summarise(
        Low = min(depth),
        Mean = mean(depth),
        High = max(depth)
      )
    summary_data
  })
}

# Run the app
shinyApp(ui, server)
```

```{r}
lake_mead_bbox <- c(left = -114.8, bottom = 36, right = -114.1, top = 36.4)
```

drought monitor shpfiles <https://droughtmonitor.unl.edu/DmData/GISData.aspx>

```{r}
install.packages("webshot")
library(leaflet)
library(htmlwidgets)
library(webshot)
kml_data <- st_read("23mead.kml")
st_write(kml_data, "23mead.shp", layer_options = "SHPT=POLYGON") 

basin <- st_read("globalwatershed.shp")

#centroid <- st_centroid(sfmead23)

#m <- leaflet(basin) %>%
# addTiles()

 m <-  leaflet() %>%
     addProviderTiles("OpenStreetMap.Mapnik") %>%
     addPolygons(data = basin, fillOpacity = 0.25)
    
    m <- addProviderTiles(m, providers$Esri.WorldImagery)
    m <- addPolygons(m, data = sfmead23)
    m
    

```

bslib package for theme

Updated app 11/29

```{r}
library(shiny)
library(leaflet)
library(htmlwidgets)
library(htmltools)
library(ggplot2)
library(dplyr)

# Define UI for application that draws a histogram
ui <- fluidPage(
  titlePanel("Lake Mead Water Depth Analysis"),
    
    mainPanel(
      tabsetPanel(
        tabPanel("Plot", plotOutput("staticPlot")), 
        tabPanel("Table", tableOutput("summaryTable"),  sliderInput("year", "Select Year:", min = 2001, max = 2022, value = 2001, step = 1, sep = "")
        ),
        tabPanel("Map", leafletOutput("map"))
      )
      )
    )

# Define server
server <- function(input, output) {

  # Create ggplot based on filtered data
  static_plot <- ggplot(mead_long, aes(x = Year, y = depth, color = depth_type)) +
    geom_line() +
    labs(title = "Yearly Water Depth for Lake Mead",
         x = "Year",
         y = "Water Depth",
         color = "Depth Type") +
    theme_minimal()
  
  # Display static plot
  output$staticPlot <- renderPlot({
    print(static_plot)
  })
  
  filtered_data <- reactive({
    filter(mead_long, Year == input$year)
  })
  
  # Create summary table based on filtered data
  output$summaryTable <- renderTable({
    summary_data <- filtered_data() %>%
      summarise(
        Low = min(depth),
        Mean = mean(depth),
        High = max(depth)
      )
    summary_data
  })

  output$map <- renderLeaflet({
    leaflet() %>%
      addProviderTiles("OpenStreetMap.Mapnik") %>%
      addPolygons(data = shapefile_2022, fillOpacity = 0.5, color = "red")
  })
}
# Run the app
shinyApp(ui, server)
```

Getting data for 2001 map (this code did not work but I want to keep this)

```

Take two

```{r}
 image_path <- sprintf("path\\to\\images\\LM%s.png", year)
```

```{r}
install.packages("bslib")
install.packages("htmltools")

library(bslib)

```

```{r}
library(sf)

basin <- st_read("globalwatershed.shp")
```

```{r}
#streamstats
library(readr)
MeadStreamStats <- read_csv("MeadStreamStats.csv")
View(MeadStreamStats)

library(readxl)
AEPmead <- read_excel("AEPmead.xlsx")
View(AEPmead)
```

AEP Curve plot

```{r}
 ggplot(AEPmead, aes(x = Statistic, y = Value)) +
  geom_line(color = "#0f4d92") +
  scale_y_continuous(labels = scales::comma) + 
  labs(
    title = "Annual Exceedance Probability Curve", 
    x = "Percent AEP Flood", 
    y = "Flow (cf/s)"
  )
```

```{r}
#Stream stats

library(readxl)
meadstats <- read_excel("MeadStats.xlsx")
```

Final app workflow

```{r}
library(shiny)
library(ggplot2)
library(dplyr)
library(magrittr)
library(imager)
library(bslib)
library(sf)
library(leaflet)
library(readxl)
library(plotly)


meaddata <- readxl::read_excel("Mead_high_low_mean.xlsx")
mead_long <- meaddata %>%
  tidyr::gather(key = "depth_type", value = "depth", -Year)
basin <- st_read("globalwatershed.shp")
AEPmead <- read_excel("AEPmead.xlsx")
meadstats <- read_excel("MeadStats.xlsx")
meadpie <- data.frame(
  Category = c("California", "Arizona", "Nevada"),
  Value = c(59, 37, 4)
 )

custom_theme <- bs_theme(
  primary = "#4D94FF",  # Set the primary color
  secondary = "#8AA9B6",  # Set the secondary color
  bg = "#E0E0E0",
  fg = "#2E4960",# Set the background color
  dark = "#4D94FF",  # Set the text color
  font_size_base = 16  # Set the base font size
)

# Define UI for application that draws a histogram
ui <- navbarPage(
  theme = custom_theme, 
  title = "Lake Mead Water Depth Analysis",
  tabPanel("Depth Data", 
           fluidRow(
             column(
               width = 6,
               p("Lake Mead, located on the Arizona-Nevada border, is a crucial reservoir that serves as a primary water source for millions of people in the southwestern United States. The lake's water is utilized for agricultural irrigation, municipal supply, and hydroelectric power generation, supporting communities in Arizona, Nevada, and California. Its geographical significance lies in its position as the largest reservoir in the United States, created by the Hoover Dam on the Colorado River.
                 However, Lake Mead is facing severe challenges, both environmental and human-induced, that have led to a significant drop in water levels. The region has been grappling with prolonged droughts, exacerbated by climate change, which contribute to decreased inflows into the reservoir. Additionally, increased water demand and the overallocation of the Colorado River's resources further strain Lake Mead.
                 Lake Mead is known for the the prominent 'bath tub rings' around the lake's perimeter. These rings, caused by the deposition of minerals and salts on exposed shorelines as the water recedes, starkly illustrate the extent of the water loss. Since the year 2000, Lake Mead has experienced a substantial decrease in water depth, losing a considerable percentage of its capacity."),
               absolutePanel(
                 fixed = TRUE, bottom = 5, left = 50,
                 width = 700, height = 300,
                 tags$img(src = "bathtubrings.jpg", width = "90%", height = "90%")
               )
             ),
             column(
               width = 6,
               plotOutput("staticPlot"),
               absolutePanel(
                tableOutput("summaryTable"),
                 fixed = TRUE, bottom = 40, right = 270,
               sliderInput("year", "Select Year:", min = 2001, max = 2022, value = 2001, step = 1, sep = "")
             ))
           )
  ),
  tabPanel("Time Lapse",
           fluidRow(
             column(
               width = 8,
           tabsetPanel(
             tabPanel("2001", tags$img(src = "LM01.png", height = 600, width = 800)),
             tabPanel("2020", tags$img(src = "LM20.png", height = 600, width = 800)))),
           column(
             width = 4,
             absolutePanel(
               fixed = TRUE, top = 175, right = 50,
               width = 525, height = 300,
             plotlyOutput("piechart")
           ))
           )),
  tabPanel("Basin",
         leafletOutput("map"),
         fluidRow(
           # Left column for the AEP plot
           column(6, 
              absolutePanel(
              fixed = FALSE, bottom = -10, left = 20,
               width = 650, height = 325, plotOutput("aepPlot"))),
           
           # Right column 
           column(6, 
             absolutePanel(
               fixed = FALSE, bottom = 20, right = -70,
               tags$style(HTML("#parameter_select { width: 300px; }
                     #parameter_table { width: 300px; }")),
                sidebarLayout(
                    sidebarPanel(
                      selectInput("parameter_select", "Select Parameter", choices = unique(meadstats$Parameter), 
                                  width = "400px"),
                      hr(),
                    ),
                    mainPanel(
                      tableOutput("parameter_table")
                    ))
                  )
                 )
                )
              )
            )
  
  
    

# Define server
server <- function(input, output) {
  
  custom_colors <- c("#191970", "#6082B6", "#337DFF")
  
  # Create ggplot based on filtered data
  static_plot <- ggplot(mead_long, aes(x = Year, y = depth, color = depth_type)) +
    geom_line() +
    theme(
      plot.background = element_rect(fill = "#D3D3D3"),  # Background color
      panel.background = element_rect(fill = "#F0F0F0")) + 
    scale_color_manual(values = custom_colors) +
    labs(title = "Yearly Water Depth for Lake Mead",
         x = "Year",
         y = "Water Depth",
         color = "Depth Type") +
    theme_minimal()
  
  # Display static plot
  output$staticPlot <- renderPlot({
    print(static_plot)
  })
  
  filtered_data <- reactive({
    dplyr::filter(mead_long, Year == input$year)
  })
  
  # Create summary table based on filtered data
  output$summaryTable <- renderTable({
    summary_data <- filtered_data() %>%
      summarise(
        Low = min(depth),
        Mean = mean(depth),
        High = max(depth)
      )
    summary_data
  })
  
  output$piechart <- renderPlotly({
    plot_ly(
      data = meadpie,
      labels = ~Category,
      values = ~Value,
      type = "pie",
      textinfo = "percent",
      hoverinfo = "label+percent",
      marker = list(colors = custom_colors)
    ) %>%
      layout(title = "Mead Water Allocation", paper_bgcolor = "#E0E0E0")
  })

  #Leaflet map of basin
  output$map <- renderLeaflet({
    leaflet() %>%
      addProviderTiles("OpenStreetMap.Mapnik") %>%
      addPolygons(data = basin, fillOpacity = 0.25)
  })
  
  #AEP curve
output$aepPlot <- renderPlot({
 ggplot(AEPmead, aes(x = Statistic, y = Value)) +
    geom_line(color = "#0f4d92") +
    scale_y_continuous(labels = scales::comma) + 
    labs(
      title = "Annual Exceedance Probability Curve", 
      x = "Percent AEP Flood", 
      y = "Flow (cf/s)"
    ) +
    theme(
      plot.margin = margin(10, 10, 10, 10, "pt"),  # Add a margin around the plot
      plot.background = element_rect(fill = NA, color = "#0f4d92", size = 1)
    )
  })
# Filter data based on selected statistic
selected_parameter_data <- reactive({
  filter(meadstats, Parameter == input$parameter_select)
})

# Display table with value and unit
observeEvent(input$parameter_select, {
  output$parameter_table <- renderTable({
    filter(meadstats, Parameter == input$parameter_select)
  })
})

# Display value and unit as text
output$selected_value_unit <- renderText({
  selected_parameter <- selected_parameter_data()
  
})
}
# Run the app
shinyApp(ui, server)
```


rsconnect::deployApp(
  appPrimaryDoc = "Lake-Mead-Water-Depth-Analysis.html",
  appFiles = c("FinalMeadApp.R", "globalwatershed.shp", "AEPmead.xlsx", "MeadStats.xlsx", "Mead_high_low_mean.xlsx", "bathtubrings.jpg", "LM01.png", "LM20.png"))
