---
title: "Bookdown"
author: "Katie Willi and Caitlin Mothes"
date: "2023-02-27"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Using `bookdown`

Put simply, R's `bookdown` package is a tool for stitching together multiple R Markdown files into a whole book. A typical `bookdown` book contains multiple *chapters*; where each *chapter* lives in its own R Markdown file. Hundreds of examples of `bookdown` documents exist, but `bookdown`'s [official website] (<https://bookdown.org/>) is a great starting point to explore some good ones.

There are several benefits of making `bookdown` documents as opposed to, say, a Microsoft Word document:

**Integration of code and text:** `bookdown` allows you to write code and text together in a single document. This makes it easy to include code examples, data visualizations, and other dynamic content alongside your written text.

**Reproducibility:** `bookdown` documents are designed to be reproducible, meaning that anyone who has access to the document can reproduce the results and figures that you've generated. This is because `bookdown` documents often include the R code that was used to generate the results, making it easy to see how the results were produced.

**Flexibility:** `bookdown` provides a lot of flexibility in terms of formatting and output options. You can create PDFs, HTML documents, e-books, and more, and you can customize the formatting and appearance of your document using a variety of built-in templates and themes.

**Collaboration:** Because `bookdown` documents are plain text files, they can easily be shared and collaborated on using GitHub. This makes it easy to work on a book with multiple authors or contributors, and to track changes and revisions over time.

**Scalability:** `bookdown` can be used to create small documents like reports and tutorials, as well as larger documents like books and dissertations. This means that `bookdown` can be used for a wide range of projects, from small data analysis projects to full-length books.

## Making a `bookdown`

In this lesson we will set up a simple `bookdown` using some of the R markdowns we have already made for this class as a starting point, then publish that document online with GitHub pages. For an example of this lesson's final outcome, see <https://kathryn-willi.github.io/my_bookdown/>.

### 1) Install `bookdown`.

```{r, warning = FALSE, eval = FALSE}
# stable version on CRAN
install.packages("bookdown")
```

### 2) Create a repository for your bookdown.

In Github, create a new repository. Call it "bookdown_demo". Clone this repository onto your local system.

### 3) Create the `index.Rmd` file.

Once you've loaded the empty repository onto your computer, create a new R Markdown file by selecting New File -\> R Markdown. Name this file **"My Bookdown"**, or whatever you fancy. This will be the title of your `bookdown` document. Save this new .Rmd file as "index.RMD" (be sure that it is saved in your **bookdown_demo** repository). This index file serves as the main entry point to your bookdown document, and will contain the YAML metadata for the entire bookdown document. Though there are many metadata options you can include for your bookdown, these are the important ones that let R know that the output for this repo is going to be a `bookdown` document, that we plan to host on GitHub:

```{r, eval = FALSE, error = FALSE}

---
title: "My Bookdown"
author: "YOUR NAME"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
description: "This is an example of using the bookdown package. The output format for this example is bookdown::gitbook."
---
  
```

### 4) Add a `bookdown` introduction to the index.

Each R Markdown file must start immediately with the chapter title using the first-level heading, e.g., `# Chapter Title`. For this first `index.Rmd` page, however, we do not want it to be labelled as a chapter. To exclude any header from being listed as a chapter, you can add `{-}` after the title. In our example and below the YAML metadata, create a non-chapter header called "My Intro", and write some introductory text below it:

```{r,  eval = FALSE, warning = FALSE, echo = TRUE}
# My Intro {-}

This is my example of a bookdown document. Blah blah blah.

```

Save then exit out of `index.RMD`.

### 5) Add chapters to your bookdown.

A typical `bookdown` book contains multiple chapters, with each chapter housed in its own R Markdown file. As mentioned previously, each R Markdown file must start immediately with the chapter title using the first-level heading (i.e. `# CHAPTER TITLE`). Only use the first-level heading at the beginning of each chapter. Let's go ahead and copy and paste each of the .Rmd files from your *Week-3-APIs-and_Iteration* into our **bookdown_demo** repository. Next, remove each of their YAML headers, and give each one a chapter name. Ensure that there are no additional first-level headers in the Rmds. Save and exit out of each RMD.

### 6) Select the order of chapters by their names.

Depending on how your *Week-3-APIs-and-Iteration* Rmds are currently named, you might need to rename them. Bookdown automatically orders chapters alphanumerically. To rename files in R, go to the Files control pane, tick the file of interest's box, then select "Rename".

### 7) Create a `_bookdown.yml` YAML file.

The `bookdown.yml` file is a configuration file used in bookdown projects. Its purpose is to define various settings and options for building the book, such as the name of the final bookdown file, where it gets saved, and other customizations. To create the `_bookdown.yml` file, go to the *Files* control pane -\> *New Blank File* -\> Text File. Name it `_bookdown.yml`. Then, copy and paste the following options into the file:

```{r, eval = FALSE, warning = FALSE, echo = TRUE}
language:
  ui:
    chapter_name: "Chapter "
output_dir: "docs" #for github pages to find it
```

### 8) "Knit" your `bookdown`.

In your **bookdown_demo** project console, run the following line of code:

```{r, eval = FALSE, warning = FALSE, echo = TRUE}
bookdown::render_book()
```

Your Rmds have now been knit, stored, and configured in a `docs` folder.

### 9) Commit and push **bookdown_demo**.

### 10) Publish your `bookdown` document on GitHub Pages.

GitHub Pages is a free web hosting service provided by GitHub that allows users to host websites directly from their GitHub repositories. These websites can be populated by a myriad of R products, including `bookdown` documents. To publish your newly-created `bookdown`, first go to your **bookdown_demo** repository on GitHub. Got to the **Settings** tab, then find **Pages**. Under the **Source** dropdown, select **Deploy from a branch**. Under the **Branch** dropdown, choose **main** and **/docs**. This tells GitHub where your `bookdown` files are located in your repo. (Remember, we explicitly listed in our YAML file that we wanted our output directory to be **docs**.)

Once these have been saved, GitHub will begin building your website. This can take a while depending on how big the website being created is. If you refresh GitHub Pages webpage, it will show the hyperlink to your new website.
